---
title: "Short Lab 0"
author: "INSERT YOUR NAME HERE"
date: "Due Date Here"
output: html_document
---

To begin, please use the following links to download R, RStudio, and R Markdown:

* [Install R](https://cran.r-project.org/)

* [Install RStudio](https://rstudio.com/products/rstudio/download/#download)

* [Install RMarkdown](https://bookdown.org/yihui/rmarkdown/installation.html#installation)

You should edit this .Rmd using RStudio, then click *Knit* in the menu bar of the Source window (above the text of this .Rmd).
Remember, you must submit *both* your .Rmd and the compiled .html in order to receive full credit!

## Part 1. Written Response

**1.** What is the name we should call you/how should we refer to you?

Replace this text with your answer.

**2.** What is your (intended) major?

Replace this text with your answer.

**3.** What is your coding/computational experience? Please specifically mention whether or not you have any experience in each of R, RMarkdown, and LaTeX. If you have used any of these tools in classes, which classes? 

Replace this text with your answer.

**4.** Please explain what you hope to take away from this course.

Replace this text with your answer.


**5.** What is the penalty if you submit an assignment 32 hours late?

Replace this text with your answer.


**6.** Do my current proposed office hour times work for you? If not, please provide some general availability.

Replace this text with your answer.

**7.** Is there anything else I should know that might help you succeed in this course? 

Replace this text with your answer.

<!--- Note: I use the backslash to skip a line for formatting. --->
\ 

## Part 2. R, RStudio, and RMarkdown

**8.** What is the relationship between R and RStudio?

**9.** Edit the code chunk below to assign the value ```5``` to the variable ```x```.

```{r}
# Your code here
```

**10.** Explain in words what the following code does. (*Optional: Why does ```print(y)``` produce 5 and not 6?*)

```{r}
y <- 5
z <- y
z <- z + 1
print(y)
print(z)
```

**11.** If you have not done so already, install the package palmerpenguins in your by running the command `install_packages("palmerpenguins")` in your console. Run the following code. 

```{r}
library(palmerpenguins)
data(penguins)
```

How many observations are in the `penguins` object? 

**12.** *Optional: What does the function defined below do? What mathematical function is this? Try running the code chunk below and then running ```h(t)``` for different values of ```t```*

```{r}
h <- function(x) {
  if (x > 1) {
    return(x * h(x - 1))
  } else {
    return(1)
  }
}
```
