---
title: "Short Lab 1"
author: "INSERT YOUR NAME HERE"
date: "Due Date Here"
output: html_document
---

<!--- Begin styling code. --->
<style type="text/css">
/* Whole document: */
body{
  font-size: 12pt;
}
h1.title {
  font-size: 38px;
  text-align: center;
}
h4.author {
  font-size: 18px;
  text-align: center;
}
h4.date {
  font-size: 18px;
  text-align: center;
}
</style>
<!--- End styling code. --->

You should edit this .Rmd using RStudio, then click *Knit* in the menu bar of the Source window (above the text of this .Rmd).
Remember, you must submit *both* your .Rmd and the compiled .html in order to receive full credit!

### Collaborators

INSERT NAMES OF ANY STUDENTS WORKED WITH

## Part 1. Some group work

Work with one or two other students, 

**1.** In words, what does the following line of code do? 

```{r, eval = T}
sample(1:6, 1)
```

**2.** Run the line of code in RStudio. What do you get? What did your groupmates get? What happens to the result each time you knit the document?

**3.** Below, we introduce the ```set.seed``` function. What do the two lines of code below do? What does the ```456``` mean? What happens to the result each time you knit the document?        

```{r, eval = T}
set.seed(456)
sample(1:6, 1)
```

**4.** We can also use in-line R code in R Markdown to ensure that we don't have to edit our text each time we knit our file. Edit the text below using in-line code to print out your value of ```x```.

```{r, eval = T}
x <- sample(1:6, 1)
```

The value of ```x``` that I obtained is YOUR ANSWER HERE.

## Part 2. Vectors and matrices

**5.** Choose a seed and replace the seed used in the code below. Using the code below, we can simulate a vector of Uniform(0,1) random variables and two additional random variables ```y``` and ```z```. What is the difference between ```y``` and ```z```?

```{r, eval = T}
set.seed(456)
x <- runif(2000)
y <- mean(x < .5)
z <- mean(x)
```

**6.** Construct a 3 by 3 matrix with 1s on the main diagonal and 3s in all other entries. Print the resulting matrix.

```{r, eval = T}
# YOUR CODE HERE
```

## Part 3. Some real data

**7.** First, run ```install.packages("moderndive")``` and then the following code. This loads the ```house_prices``` dataset, which contains house sale data for King Country. What time period are these house prices from?

```{r}
library(moderndive)
data("house_prices")
```

**8.** The ```house_prices$bedrooms``` vector contains information on the number of bedrooms in each house. What is the median value? What is the most common value?

```{r}
# Your code here
```

**9.** (Optional) The ```house_prices$bedrooms``` vector contains information on the number of bedrooms in each house. What is the maximum value? Is this realistic? Can you figure out what happened here?