---
title: "Lab 4"
author: "STAT 302"
date: "Due Date Here"
output: html_document
---

<!--- Begin styling code. --->
<style type="text/css">
/* Whole document: */
body{
  font-family: "Palatino Linotype", "Book Antiqua", Palatino, serif;
  font-size: 12pt;
}
h1.title {
  font-size: 38px;
  text-align: center;
}
h4.author {
  font-size: 18px;
  text-align: center;
}
h4.date {
  font-size: 18px;
  text-align: center;
}
</style>
<!--- End styling code. --->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*If you collaborated with anyone, you must include "Collaborated with: FIRSTNAME LASTNAME" at the top of your lab!*

## Part 1. Training and Test Error (10 points)

Use the following code to generate data:

```{r, message = FALSE}
library(ggplot2)
# generate data
set.seed(302)
n <- 30
x <- sort(runif(n, -3, 3))
y <- 2*x + 2*rnorm(n)
x_test <- sort(runif(n, -3, 3))
y_test <- 2*x_test + 2*rnorm(n)
df_train <- data.frame("x" = x, "y" = y)
df_test <- data.frame("x" = x_test, "y" = y_test)

# store a theme
my_theme <- theme_bw(base_size = 16) + 
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5))

# generate plots
g_train <- ggplot(df_train, aes(x = x, y = y)) + geom_point() +
  xlim(-3, 3) + ylim(min(y, y_test), max(y, y_test)) + 
  labs(title = "Training Data") + my_theme
g_test <- ggplot(df_test, aes(x = x, y = y)) + geom_point() +
  xlim(-3, 3) + ylim(min(y, y_test), max(y, y_test)) + 
  labs(title = "Test Data") + my_theme
g_train
g_test
```

**1a.** For every k in between 1 and 10, fit a degree-k polynomial linear regression model with `y` as the response and `x` as the explanatory variable(s).
(*Hint: Use *`poly()`*, as in the lecture slides.*)

**1b.** For each model from (a), record the training error. Then predict `y_test` using `x_test` and also record the test error.
(*Hint: You can either do this part by copying your code 10 times for each value of k (which is totally fine!), or in a loop (if you are struggling to call your model within a loop check out the functions get() and paste0()*)

**1c.** Present the 10 values for both training error and test error on a single table. Comment on what you notice about the relative magnitudes of training and test error, as well as the trends in both types of error as $k$ increases.

**1d.** If you were going to choose a model based on training error, which would you choose? Plot the data, colored by split. Add a line to the plot representing your selection for model fit. Add a subtitle to this plot with the (rounded!) test error.
(*Hint: See Lecture Slides 8 for example code.*)

**1e.** If you were going to choose a model based on test error, which would you choose? Plot the data, colored by split. Add a line to the plot representing your selection for model fit. Add a subtitle to this plot with the (rounded!) test error.

**1f.** What do you notice about the shape of the curves from part (d) and (e)? Which model do you think has lower bias? Lower variance? Why?

## Part 2. k-Nearest Neighbors Cross-Validation (10 points)

For this part, note that there are tidyverse methods to perform cross-validation in R (see the `rsample` package). However, your goal is to understand and be able to implement the algorithm "by hand", meaning  that automated procedures from the `rsample` package, or similar packages, will not be accepted.

To begin, load in the popular `penguins` data set from the package `palmerpenguins`.

```{r}
library(palmerpenguins)
data("penguins")
```

Our goal here is to predict output class `species` using covariates `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, and `body_mass_g`.
All your code should be within a function `my_knn_cv`.

**Input:**

  * `train`: input data frame
  * `cl`: true class value of your training data
  * `k_nn`: integer representing the number of neighbors
  * `k_cv`: integer representing the number of folds
  
*Please note the distinction between `k_nn` and `k_cv`!*

**Output:** a list with objects

  * `class`: a vector of the predicted class $\hat{Y}_{i}$ for all observations
  * `cv_err`: a numeric with the cross-validation misclassification error


You will need to include the following steps:

* Within your function, define a variable `fold` that randomly assigns observations to folds $1,\ldots,k$ with equal probability. (*Hint: see the example code on the slides for k-fold cross validation*)
* Iterate through $i = 1:k$. 
  * Within each iteration, use `knn()` from the `class` package to predict the class of the $i$th fold using all other folds as the training data.
  * Also within each iteration, record the misclassification rate (a value between 0 and 1 representing the proportion of observations that were classified **incorrectly**).
* After you have done the above steps for all $k$ iterations, store the vector `class` as the output of `knn()` with the full data as both the training and the test data, and the value `cv_error` as the average misclassification rate from your cross validation.

**Submission:** To prove your function works, apply it to the `penguins` data. Predict output class `species` using covariates `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, and `body_mass_g`. You will need to first filter out any observations that have missing values for any of these covariates. Use $5$-fold cross validation (`k_cv = 5`). Use a table to show the `cv_err` values for 1-nearest neighbor and 5-nearest neighbors (`k_nn = 1` and `k_nn = 5`), as well as the training set error values for 1 and 5 nearest neighbors (compare your output `class` to the true class, `penguins$species`). Comment on which value had lower CV misclassification error and which had lower training set error.